A Simplified Plackett-Luce Likelihood Function for Rank Orders

### 2025-10-17 inst/exponential_BT.Rmd
more work

```{r allthree}
worker_allthree <- function(...){
op <- par(mfrow=c(1,3))
worker_123first()
worker_123second()
worker_123third()
dev.off()
par(op)
}
worker_allthree()
pdf(file="123all.pdf",height=4)
worker_allthree()
dev.off()
```
The $n=4$ case is harder as we potentially have six consistent
orderings for each observation.  As an example, if the focal
competitor comes third we have $1$

>>>>

```{r allthree}
worker_allthree <- function(...){
op <- par(mfrow=c(1,3))
worker_123first()
worker_123second()
worker_123third()
dev.off()
par(op)
}
worker_allthree()
pdf(file="123all.pdf",height=4)
worker_allthree()
dev.off()
```
The $n=4$ case is harder as we potentially have six consistent
orderings for each observation.  As an example, if the focal
competitor comes third we have

\begin{equation}
\newcommand{pl}[4]{
\frac{x^{#1}}{x^{#1} + x^{#2} + x^{#3} + x^{#4}}\cdot
\frac{x^{#2}}{         x^{#2} + x^{#3} + x^{#4}}\cdot
\frac{x^{#3}}{                  x^{#3} + x^{#4}}}

\end{equation}

SDAdsfDSFFDSAfdsafdsff asdf dfs dfs dfs dsff dsfds fds fds sfd s fs sd sdf sdf sdf sdf sdf sdf sdf sdfs dfsdf sdsdfsa asdf saf

\begin{eqnarray}
\mathcal{L}(x;r=1)
\quad &=& \quad\hphantom{+}
            \pl{2}{3}{1}{4}\\
&{}&\quad + \pl{2}{4}{1}{3}\\
&{}&\quad + \pl{3}{2}{1}{4}\\
&{}&\quad + \pl{3}{4}{1}{2}\\
&{}&\quad + \pl{4}{2}{1}{3}\\
&{}&\quad + \pl{4}{3}{1}{2}
\end{eqnarray}

\begin{eqnarray}
\mathcal{L}(x;r=2) 
\quad &=& \quad\hphantom{+}
            \pl{1}{3}{2}{4}\\
&{}&\quad + \pl{1}{4}{2}{3}\\
&{}&\quad + \pl{3}{1}{2}{4}\\
&{}&\quad + \pl{3}{4}{2}{1}\\
&{}&\quad + \pl{4}{1}{2}{3}\\
&{}&\quad + \pl{4}{3}{2}{1}
\end{eqnarray}

and so on.

```{r fourcompetitors}
worker <- function(...){
PL <- function(v){
  stopifnot(length(v) == 4)
  v[1]/(v[1]+v[2]+v[3]+v[4]) *
  v[2]/(     v[2]+v[3]+v[4]) *
  v[3]/(          v[3]+v[4])
}

f1 <- function(v){  # focal competitor came third, r=1
  return(
  PL(v^c(2,3,1,4)) +
  PL(v^c(2,4,1,3)) +
  PL(v^c(3,2,1,4)) +
  PL(v^c(3,4,1,2)) +
  PL(v^c(4,2,1,3)) +
  PL(v^c(4,3,1,2))
  )
}
f2 <- function(v){  # focal competitor came third, r=2
  return(
  PL(v^c(1,3,2,4)) +
  PL(v^c(1,4,2,3)) +
  PL(v^c(3,1,2,4)) +
  PL(v^c(3,4,2,1)) +
  PL(v^c(4,1,2,3)) +
  PL(v^c(4,3,2,1))
  )
}
f3 <- function(v){  # focal competitor came third, r=2
  return(
  PL(v^c(1,2,3,4)) +
  PL(v^c(1,4,3,2)) +
  PL(v^c(2,1,3,4)) +
  PL(v^c(2,4,3,1)) +
  PL(v^c(4,1,3,2)) +
  PL(v^c(4,2,3,1))
  )
}
f4 <- function(v){  # focal competitor came third, r=4
  return(
  PL(v^c(1,2,4,3)) +
  PL(v^c(1,3,4,2)) +
  PL(v^c(2,1,4,3)) +
  PL(v^c(2,3,4,1)) +
  PL(v^c(3,1,4,2)) +
  PL(v^c(3,2,4,1))
  )
}

o <- 1:4

plot(c(-3,0),c(-3,0), type='n',
     xlab="log x", ylab="log-likelihood", main="Focal competitor came third")
x <- seq(from=0.01,to=1,by=0.001)

points(log(x),log(sapply(x,f1)), type="l", col="black")
points(log(x),log(sapply(x,f2)), type="l", col="red")
points(log(x),log(sapply(x,f3)), type="l", col="blue")
points(log(x),log(sapply(x,f4)), type="l", col="green")

abline(h=0,lty=2)
abline(h = -2, lty=1)

legend("topright",lty=1,col=c("black","red","blue","green"),
       legend=c("r=1","r=2","r=3","r=4"), bg="white")
}

worker()
pdf(file="1234third.pdf")
worker()
dev.off()
```

# More stuff

### 2025-10-17 inst/hankin-exponential_BT.tex
more work

The $n=4$ case is harder as we potentially have six consistent
orderings for each observation.  As an example, if the focal
competitor comes third we have $1$

>>>>

The $n=4$ case is harder as we potentially have six consistent
orderings for each observation.  As an example, if the focal
competitor comes third we have

The $n=4$ case is harder as we potentially have six consistent
orderings for each observation.  As an example, if the focal
competitor comes third we have

SDAdsfDSFFDSAfdsafdsff asdf dfs dfs dfs dsff dsfds fds fds sfd s fs sd
sdf sdf sdf sdf sdf sdf sdf sdfs dfsdf sdsdfsa asdf saf

\newcommand{\plackett}[4]{
\frac{x^{#1}}{x^{#1} + x^{#2} + x^{#3} + x^{#4}}\cdot
\frac{x^{#2}}{         x^{#2} + x^{#3} + x^{#4}}\cdot
\frac{x^{#3}}{                  x^{#3} + x^{#4}}}

\begin{eqnarray}
\mathcal{L}(x;r=1)
\quad &=& \quad\hphantom{+}
            \plackett{2}{3}{1}{4}\\
&{}&\quad + \plackett{2}{4}{1}{3}\\
&{}&\quad + \plackett{3}{2}{1}{4}\\
&{}&\quad + \plackett{3}{4}{1}{2}\\
&{}&\quad + \plackett{4}{2}{1}{3}\\
&{}&\quad + \plackett{4}{3}{1}{2}
\end{eqnarray}

%\begin{eqnarray}
%\mathcal{L}(x;r=2) 
%\quad &=& \quad\hphantom{+}
%            \pl{1}{3}{2}{4}\\
%&{}&\quad + \pl{1}{4}{2}{3}\\
%&{}&\quad + \pl{3}{1}{2}{4}\\
%&{}&\quad + \pl{3}{4}{2}{1}\\
%&{}&\quad + \pl{4}{1}{2}{3}\\
%&{}&\quad + \pl{4}{3}{2}{1}
%\end{eqnarray}

and so on.

\begin{figure}[t]
  \begin{centering}
\includegraphics[width=4in]{1234third}  % 1234third.pdf produced by
                                       % very_simplified_likelihood.Rmd
\caption{Likelihood function for the observation that the focal
  competitor came third out of four, showing
  $r=1,2,3,4$\label{1234third}.  Value of $r$ indicated in colour by
  its curve; dotted line shows (asymptotic) maximum likelihood
  estimate (viz, $r=3, x\longrightarrow 0$)}
\end{centering}
\end{figure}
